%! TeX program = lualatex
\documentclass[a4paper]{article} 

% packages
\usepackage{microtype}      % Slightly tweak font spacing for aesthetics
\usepackage[english]{babel} % Language hyphenation and typographical rules
\usepackage[final, colorlinks = false, urlcolor = cyan]{hyperref} 
\usepackage{changepage}     % adjust margins on the fly

\usepackage{fontspec}
\setmainfont{EB Garamond}
\setmonofont[Scale=MatchLowercase]{Deja Vu Sans Mono}

\usepackage{minted}
\usemintedstyle{algol_nu}
\usepackage{xcolor}

\usepackage{pgfplots}
\pgfplotsset{width=\textwidth,compat=1.9}

\usepackage{caption}
\newenvironment{code}{\captionsetup{type=listing}}{}
\captionsetup[listing]{skip=0pt}
\setlength{\abovecaptionskip}{5pt}
\setlength{\belowcaptionskip}{5pt}

\usepackage[yyyymmdd]{datetime}
\renewcommand{\dateseparator}{--}

\usepackage{titlesec}
% \titleformat{\section}{\LARGE\bfseries}{}{}{}[\titlerule]
% \titleformat{\subsection}{\Large\bfseries}{}{0em}{}
% \titlespacing{\subsection}{0em}{-0.7em}{0em}
%
% \titleformat{\subsubsection}{\large\bfseries}{}{0em}{$\bullet$ }
% \titlespacing{\subsubsection}{1em}{-0.7em}{0em}

% margins
\addtolength{\hoffset}{-2.25cm}
\addtolength{\textwidth}{4.5cm}
\addtolength{\voffset}{-3.25cm}
\addtolength{\textheight}{5cm}
\setlength{\parskip}{0pt}
\setlength{\parindent}{0in}
% \setcounter{secnumdepth}{0}

\begin{document}
\hrule \medskip
\begin{minipage}{0.295\textwidth} 
    \raggedright
    \footnotesize 
    Name: Andrew Hayes \\
    E-mail: \href{mailto://a.hayes18@universityofgalway.ie}{a.hayes18@universityofgalway.ie}  \hfill\\   
    ID: 21321503 \hfill
\end{minipage}
\begin{minipage}{0.4\textwidth} 
    \centering 
    \vspace{0.4em}
    \Large 
    \textbf{CT3112} \\ 
\end{minipage}
\begin{minipage}{0.295\textwidth} 
    \raggedleft
    \today
\end{minipage}
\medskip\hrule 
\begin{center}
    \normalsize
    Assignment 01: Presentation Skills
\end{center}
\hrule

\begin{itemize}
    \item   The role of the presenter.
    \item   Preparation \& Planning: Why, Who, Where, When, What, \& How.
    \item   Structuring the content for maximum impact:
            \begin{itemize}
                \item   Introduction to the topic.
                \item   The middle part.
                        \begin{itemize}
                            \item   Aligned with the introduction.
                            \item   Main topics.
                            \item   Each main topic has sub-topics.
                            \item   Logical sequence.
                                    \begin{itemize}
                                        \item   What are the main points I need to address?
                                        \item   What does the audience need to know first/second etc?
                                        \item   What do they need to know by the end?
                                        \item   Are my main points structured in a logical/coherent way?
                                        \item   What evidence will I use to support these points?
                                    \end{itemize}
                        \end{itemize}
                \item   The conclusion.
            \end{itemize}

            The presentation should be written in the following order:
            \begin{enumerate}
                \item   Purpose \& Objective.
                \item   Middle: main content.
                \item   End: Summary, conclusions, \& recommendations.
                \item   Beginning: Introduction/Opening.
            \end{enumerate}

            The content should be logical \& understandable. Use storyboards/maps.
            Go from simple to complex, presented incrementally, concisely, \& in order.
            Use chronological ordering.
            Have clear sections \& headings to provide structure.
            Lead from one point to another.

    \item   Effective delivery techniques.
            \begin{itemize}
                \item   Clarity.
                \item   Brevity.
                \item   Concision.
            \end{itemize}
    \item   Stages of the communication process
            \begin{itemize}
                \item   Encoding.
                \item   Transmission.
                \item   Decoding.
                \item   Feedback.
            \end{itemize}
    \item   Barriers to effective communication:
            \begin{itemize}
                \item   Language barriers.
                \item   Cultural barriers.
                \item   Technology barriers.
            \end{itemize}
\end{itemize}



\end{document}
